Supplementary Materials

### 2023-05-17 analysis/supplementary-materials/supp-mat.qmd
clean analysis code

# upload data
metadata <- read_tsv(here("analysis/data/raw_data/metadata.tsv"))
demography <- read_csv(here("analysis/data/raw_data/demography.csv"))
lloq <- read_tsv(here("analysis/data/raw_data/lloq.tsv"))
uhplc_data_comb <- read_csv(here("analysis/data/derived_data/uhplc-data_combined.csv"))
#uhplc_calculus <- read_csv(here("analysis/data/derived_data/uhplc-calculus_cleaned.csv"))
dental_inv <- read_csv(here("analysis/data/raw_data/dental-inv.csv"))
caries <- read_csv(here("analysis/data/raw_data/caries.csv"))
periodont <- read_csv(here("analysis/data/raw_data/periodontitis.csv"))
periap <- read_csv(here("analysis/data/raw_data/periapical.csv"))
calculus <- read_csv(here("analysis/data/raw_data/calculus.csv"))
calculus_full <- read_csv(here("analysis/data/raw_data/calculus_full.csv"))
sinusitis_clean <- read_csv(here("analysis/data/derived_data/sinusitis_cleaned.csv"))
path_cond_clean <- read_csv(here("analysis/data/derived_data/path-conditions_cleaned.csv"))

@@

All raw data are available for download from Zenodo.

## Sample

@@

```{r}
#| label: tbl-sampling
#| tbl-cap: "Table showing which individuals were sampled in this study and which individuals were sampled in the previous study. When both TRUE, the individual was sampled twice."
metadata %>% 
  select(id, element, KZ_element) %>% 
  mutate(
    element = case_when(is.na(element) ~ FALSE,
                   TRUE ~ TRUE),
    KZ_element = case_when(is.na(KZ_element) ~ FALSE,
                   TRUE ~ TRUE),
    ) %>%
  knitr::kable(col.names = c("ID", "this study", "previous study"))
```

@@

```{r}
tobacco <- uhplc_calculus_long %>%
  filter(compound %in% c("nicotine", "cotinine")) %>% 
  mutate(detection = if_else(is.na(conc), FALSE, TRUE))

@@

# accuracy in replicated samples
uhplc_accuracy <- uhplc_calculus_replicated %>%
  left_join(select(demography, id, pipe_notch), by = "id") %>% 
  filter(compound == "nicotine" | compound == "cotinine") %>%
  group_by(id) %>% 
  arrange(desc(presence)) %>%
  slice(which(compound == "nicotine" || compound == "cotinine")) %>%
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(presence == 1 & pipe_notch == "Y" ~ TRUE,
                             presence == 0 & pipe_notch == "N" ~ TRUE, # can we be sure if this is correct?
                             presence == 1 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
                             TRUE ~ FALSE))

replicated_accuracy <-

@@

>>>>

# upload data
# metadata <- read_tsv(here("analysis/data/raw_data/metadata.tsv"))
# demography <- read_csv(here("analysis/data/raw_data/demography.csv"))
# lloq <- read_tsv(here("analysis/data/raw_data/lloq.tsv"))
# uhplc_data_comb <- read_csv(here("analysis/data/derived_data/uhplc-data_combined.csv"))
# #uhplc_calculus <- read_csv(here("analysis/data/derived_data/uhplc-calculus_cleaned.csv"))
# dental_inv <- read_csv(here("analysis/data/raw_data/dental-inv.csv"))
# caries <- read_csv(here("analysis/data/raw_data/caries.csv"))
# periodont <- read_csv(here("analysis/data/raw_data/periodontitis.csv"))
# periap <- read_csv(here("analysis/data/raw_data/periapical.csv"))
# calculus <- read_csv(here("analysis/data/raw_data/calculus.csv"))
# calculus_full <- read_csv(here("analysis/data/raw_data/calculus_full.csv"))
# sinusitis_clean <- read_csv(here("analysis/data/derived_data/sinusitis_cleaned.csv"))
# path_cond_clean <- read_csv(here("analysis/data/derived_data/path-conditions_cleaned.csv"))

@@

## Samples

@@

```{r}
#| label: tbl-sampling
#| tbl-cap: "Table showing which individuals were sampled in this study and which individuals were sampled in the previous study. When both TRUE, the individual was sampled twice."
metadata %>% 
  select(id, element, KZ_element) %>% 
  mutate(
    element = case_when(is.na(element) ~ FALSE,
                   TRUE ~ TRUE),
    KZ_element = case_when(is.na(KZ_element) ~ FALSE,
                   TRUE ~ TRUE),
    ) %>%
  arrange(id) %>% 
  knitr::kable(col.names = c("ID", "this study", "previous study"))
```

@@

```{r}
#| label: tobacco-accuracy-setup
tobacco <- uhplc_calculus_long %>%
  filter(compound %in% c("nicotine", "cotinine")) %>% 
  mutate(detection = if_else(is.na(conc), FALSE, TRUE))

@@

# accuracy in replicated samples
uhplc_accuracy <- uhplc_calculus_replicated %>%
  left_join(select(demography, id, pipe_notch), by = "id") %>% 
  filter(compound == "nicotine" | compound == "cotinine") %>%
  group_by(id) %>% 
  arrange(desc(presence), .by_group = T) %>%
  distinct(id, .keep_all = T) %>% 
  #slice(which(str_detect(compound, "nicotine|cotinine"))) #%>%
  # slice(which(compound == "nicotine" | compound == "cotinine")) %>%
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(presence == 1 & pipe_notch == "Y" ~ TRUE,
                             presence == 0 & pipe_notch == "N" ~ TRUE, # can we be sure if this is correct?
                             presence == 1 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
                             TRUE ~ FALSE))

@@

All raw data are available for download from Zenodo (<https://doi.org/10.5281/zenodo.7648757>).

### lloq.tsv

| variable | description |
|---|---|
| compound | name of target compound |
| lloq | Lower limit of quantitation |
